---
title: "analysis-UniqueFeederCount-Age-Sex"
author: "Margaret Dougher"
format: html
editor: visual
---

In this file I am going to be performing a statistical analysis test, Chi-square test of independence, to look for a relationship between Age and Sex, and Unique Feeder Counts of Black chickadees.

## Set up
```{r}
rm(list=ls()) #clear's R's brain
library(here)
library(tidyverse)
library(ggfortify)
```

## Reading in the data
```{r}
dates <- read.csv(here("Data", "dates.csv"))
```

Question: Is there an association between the Age and/or Sex and the Unique Feeder Count of Black chickadees?

Prediction: I believe that Age will be associated with Unique Feeder Count, but Sex will not.

I am going to use group_by and summarise to organize the data I will be working with.

```{r}
totals <- dates |>
  group_by(Age, Sex) |>
  summarise(total.uniquefeedercount = sum(UniqueFeederCount))
```

After doing this I looked at the table and noticed that I did not have much data for Unique Feeder Counts because a lot of them in the table are NA's. Is there another way I should be grouping the data?

I have decided to remove the NA's and just work with complete cases in my data.

```{r}
df_remove_All <- totals[complete.cases(totals), ]
```

Since I am working with count data I am going to use a bar graph to plot the data.

## Plotting the data

```{r}
ggplot(df_remove_All, aes(x = Age, y = total.uniquefeedercount, fill = Sex))+
     geom_bar(stat = "identity", position = "dodge")
```

Looking at this graph I think that Sex has more of an association to Unique Feeder Count than Age. I would expect to reject my null hypothesis that Age will have an association with Unique Feeder Count and Sex will not.

Now I need to create a matrix in order to perform the chi-square test of independence.

```{r}
dates.mat <- xtabs(total.uniquefeedercount ~ Age + Sex, data = df_remove_All)
```

Now that I have created the matrix I can run the test.

```{r}
chisq.test(dates.mat)
```

## Interpreting my results
There is a very small chance that we are seeing more of an association betwen Age and Unique Feeder Count. Therefore I can reject my null hypothesis that Age has an association with Unique Feeder Count, and Sex does not (X-squared = 202.07, df = 3, p-value = < 2.2e-16).

The last step is to interpret what the test shows me, but I am having a problem with my matrix it says that all entries of x must be nonnegative for the chi-square test, but I dont have any negative values in my data. I am unsure of how to fix this.

I think I solved the problem by only using complete cases but I am still unsure.