---
title: "Analysis-"
author: "Margaret Dougher"
format: html
editor: visual
---

In this file I am going to be performing a statistical analysis test, linear regression, to look for a relationship between AvgTemp_c and VisitCount of Black chickadees.

## Set up

```{r}
rm(list=ls()) #clear's R's brain
library(here)
library(tidyverse)
library(ggfortify)
```

## Reading in the data

```{r}
dates <- read.csv(here("Data", "dates.csv"))
```

Question: Is there a relationship between the AvgTemp_C and the VisitCount of Black chickadees?

Prediction: The lower the average temperature, the less times Black chickadees with visit a feeder.

## Plotting the data

Starting with a scatterplot

```{r}
ggplot(dates, aes(x = AvgTemp_C, y = VisitCount)) +
  geom_point(size = 1) +
  theme_bw()
```

## Interpreting the figure

The slope is relatively positive, and roughly speaking, the VisitCount ranges between **0-280**. The AvgTemp_c ranges between **-18 to 5**. Looking at the graph I can also see that the y-intercept is somewhere around 0.

## Making a simple linear regression

```{r}
model_dates <- lm(VisitCount ~ AvgTemp_C, data = dates)
```

After making the model I am now going to look at some assumptions.

```{r}
autoplot(model_dates)
```

## Making an interpretation

```{r}
anova(model_dates)
```

This table shows me that the F-value is rather large, indicating that there is most likely a significant difference between AvgTemp_C and VisitCount.

I am also looking at a summary table for further analysis.

```{r}
summary(model_dates)
```

This shows us that the p-value is very small meaning there is a significant difference between AvgTemp_C and VisitCount. It also gives us our degrees of freedom, and our F-value. I will take into account that 171 observations were not included in this summary due to them being NA's.

## Final plot

Now I am going to make a figure that shows off the relationship of my statistical results.

```{r}
ggplot(dates, aes(x = AvgTemp_C, y = VisitCount)) +
  geom_point(size = 1) +
  geom_smooth(method = 'lm') + #added the fitted values and standard error to the figure
  ylab("Visit Count") + #alters label of y and x axis
  xlab("Average Temperature (C)") +
  theme_bw()
```

I am unsure if this is the proper graph I should use, and if adding the geom_smooth() into the picture was the right idea?
