---
title: "Analysis-"
author: "Margaret Dougher"
format: html
editor: visual
---

In this file I am going to be performing a statistical analysis test, linear regression, to look for a relationship between AvgTemp_c and VisitCount of Black chickadees.

## Set up

```{r}
rm(list=ls()) #clear's R's brain
library(here)
library(tidyverse)
library(ggfortify)
```

## Reading in the data

```{r}
dates <- read.csv(here("Data", "dates.csv"))
```

Question: Is there a relationship between the AvgTemp_C and the VisitCount of Black chickadees?

Prediction: The lower the average temperature, the more times Black chickadees will visit a feeder. This is because the colder it gets outside the more energy the chickadees will need, and they get their energy from consuming food.

## Plotting the data

Starting with a scatterplot

```{r}
ggplot(dates, aes(x = AvgTemp_C, y = VisitCount)) +
  geom_point(size = 1) +
  xlab("Visit Count") +
  ylab("Average Temperature_C")
  theme_bw()
```

## Interpreting the figure

The slope is relatively positive, and roughly speaking, the VisitCount ranges between 0-280. The AvgTemp_c ranges between -25 to 5. Looking at the graph I can also see that the y-intercept is somewhere around 50. I think that the slope (rise/run) will be around 14. 

## Making a simple linear regression

```{r}
model_dates <- lm(VisitCount ~ AvgTemp_C, data = dates)
```

After making the model I am now going to look at some assumptions.

```{r}
autoplot(model_dates)
```

This is showing me that I am performing the right test, linear regression, and that my data is pretty linear, but the normality looks a little bit off.

I am going to produce a histogram to look more closely at my data.

```{r}
ggplot(dates, aes(x = VisitCount)) +
  geom_histogram()
```

From this graph I can see that my values are slightly right skewed, which would explain why my normality is a bit off. 

Because of this I am going to be performing a a GLM test.

```{r}
model_dates_1 <- glm(VisitCount ~ AvgTemp_C, data = dates, family = poisson)
```

After creating my new model using glm, I am going to now check my assumptions again with my new formatted data.

```{r}
autoplot(model_dates_1)
```

I can see now that although my normal distribution is not perfect, using the glm reduced some of the abnormality. 

Now that my data is better fitted, I can start performing my statistical analysis.

```{r}
summary(model_dates_1)
anova(model_dates_1, test = "Chisq")
```

Although temperature is statistically meaning full according to my p-value (p = 2e -16), it is not explaining a lot of the variance in the data. This is shown because the null deviance (102485) and the residual deviance (101728) are so close. 

Now I am going to make a figure that shows off the relationship of my statistical results, but first I need to alter some of my data sets and make some new ones in order to plot the proper final graph.

```{r}
min.temp <- min(dates$AvgTemp_C)
max.temp <- max(dates$AvgTemp_C)
```

```{r}
new.x <- expand.grid(AvgTemp_C = seq(min.temp, max.temp, length = 1000)) #creating the new.x values using the variable AvgTemp_C
```

```{r}
# generating fits and standard errors at new.x values
new.y <- predict(model_dates_1, newdata = new.x, se.fit = T)
new.y <- data.frame(new.y)
```

```{r}
addThese <- data.frame(new.x, new.y) 
```

```{r}
# adding the confidence intervals into the new data frame
addThese <- mutate(addThese,
  VisitCount = exp(fit), 
  lwr = exp(fit - 1.96 * se.fit),
  upr = exp(fit + 1.96 * se.fit)
)
```

## Creating the final plot

```{r}
ggplot(dates, aes(AvgTemp_C, VisitCount)) +
  geom_point(size = 2) +
  geom_smooth(data = addThese, aes(ymin = lwr, ymax = upr), stat = "identity")
```

## Results Statement
 
