---
title: "Dougher-Capstone-Project"
author: "Margaret Dougher"
format: html
editor: visual
---

## Setting things up

```{r}
rm(list=ls()) #clears R's brain
library(here) #lets R know where the file path is starting
library(tidyverse) #inserting necessary packages
library(ggfortify)
library(multcomp)
```

## Importing Data Sets

```{r}
feeder_visits <- read.csv(here("Data", "data_FeederVisits (1).csv"))
survival <- read.csv(here("Data", "THC_Survival (1).csv"))
dates <- read.csv(here("Data", "data_AllDates (2).csv"))
```

## Examining my data

```{r}
glimpse(dates) #looking at data
```

I can see from using the glimpse function that for the data set dates, I have 10 columns of different variables, and I have 5,796 rows of different data. Looking at my data I need to change VisitDate from a character variable to a date variable. I also need to change Age_Sex and Sex from character variables to factor variables.

## Changing variables

```{r}
dates$Age_Sex <- as.factor(dates$Age_Sex) #changing to factor variable
dates$Sex <- as.factor(dates$Sex) 
dates$VisitDate <- mdy(dates$VisitDate) #changing to date variable

levels(dates$Age_Sex) #looking at different levels of Age_Sex
levels(dates$Sex)
```

## Examining my data

```{r}
glimpse(feeder_visits) #looking at data
```

Looking at my data I can see that there are 6 columns of different variables of data, and there is 392,974 rows of data. I do need to change some of variables data types. I need to change VisitDate to a date data type not a character variable. I also need to change Feeder from a character variable to a factor variable.

## Changing variables

```{r}
feeder_visits$VisitDate <- mdy(feeder_visits$VisitDate) #changing to date variable
feeder_visits$Feeder <- as.factor(feeder_visits$Feeder) #changing to factor variable

levels(feeder_visits$Feeder) #looking at different levels of feeders
```

## Examining my data

```{r}
glimpse(survival) 
```

From using the glimpse function I can see that my data has 2 columns, TransponderHexCode, and Survived. It also shows 138 rows of data. I do not need to change any of the data types for either of the variables.

## Looking for outliers using with the data and variables I will be looking at for relationships

```{r}
# plotting the data to look for outliers using box plots

ggplot(dates, aes(x = AvgTemp_C, y = VisitCount)) +
  geom_boxplot() +
  xlab("Average Temperature (C)") +
  ylab("Visit Count") +
  coord_flip() +
  theme_bw()

ggplot(dates, aes(x = Age, y = UniqueFeederCount)) +
  geom_boxplot() +
  theme_bw()

ggplot(dates, aes(x = Sex, y = UniqueFeederCount)) +
  geom_boxplot() +
  theme_bw()
```

I am going to examine my data using some histogram plots to get a different view of my data 

```{r}
ggplot(dates, aes(VisitCount))+
  geom_histogram()

ggplot(dates, aes(UniqueFeederCount))+
  geom_histogram()
```

Now I am going to use the arrange function to further look for outliers

```{r}
dates %>% arrange(AvgTemp_C) %>% head(10) %>% View() #shows upper half of data
dates %>% arrange(AvgTemp_C) %>% tail(10) %>% View() #shows lower half of data

dates %>% arrange(VisitCount) %>% head(10) %>% View()
dates %>% arrange(VisitCount) %>% tail(10) %>% View()

dates %>% arrange(Age) %>% head(10) %>% View()
dates %>% arrange(Age) %>% tail(10) %>% View()

dates %>% arrange(Sex) %>% head(10) %>% View()
dates %>% arrange(Sex) %>% tail(10) %>% View()

dates %>% arrange(UniqueFeederCount) %>% head(10) %>% View()
dates %>% arrange(UniqueFeederCount) %>% tail(10) %>% View()
```

In looking for outliers I realized that I have a few outliers in my data, but none that are prevelant enough for me to remove or alter in anyway. I have decided to leave my data the way it is.

## Saving clean and revised data

```{r}
write.csv(dates,here("Data", "dates.csv"), row.names = FALSE) #row.names=FALSE prevents R from adding a column of row numbers to the output file

write.csv(feeder_visits,here("Data", "feeder_visits.csv"), row.names = FALSE)

write.csv(survival,here("Data", "survival.csv"), row.names = FALSE)
```

## Looking at the data using exploratory data anaylsis

```{r}
summary(dates)

dates |>
  group_by(AvgTemp_C) |>
  summarise(VisitCount) #looking at just these two variables together

dates |>
  group_by(AvgTemp_C) |>
  summarise(meanVisitCount = mean(VisitCount)) #looking at the mean number of visits at each temperature

dates |>
  group_by(Age) |>
  summarise(UniqueFeederCount)

dates |>
  group_by(Age) |>
  summarise(meanUniqueFeederCount = mean(UniqueFeederCount)) #looking at the mean number of unique feeder visits at each age

dates |>
  group_by(Sex) |>
  summarise(UniqueFeederCount)

dates|> 
  group_by(Sex) |>
  summarise(meanUniqueFeederCount = mean(UniqueFeederCount)) #looking at the mean number of unique feeder visits by each sex
```

By using the group by and summarise function for my various data, I was able to determine the best way to isolate my data for analysis and for me to view and understand.

## Plotting my data for analysis

```{r}
ggplot(dates, aes(x = AvgTemp_C, y = VisitCount)) +
  geom_boxplot() +
  xlab("Average Temperature (C)") +
  ylab("Visit Count") +
  theme_bw()

ggplot(dates, aes(x = AvgTemp_C, y = VisitCount)) +
  geom_point(size = 1) +
    xlab("Average Temperature (C)") +
    ylab("Visit Count") +
    theme_bw()
```

## Continuing to plot my data for analysis

```{r}
ggplot(dates, aes(x = Age, y = UniqueFeederCount)) +
  geom_boxplot() +
  theme_bw()

ggplot(dates, aes(x = Age, y = UniqueFeederCount)) +
  geom_point(size = 1) +
    theme_bw()

ggplot(dates, aes(x = Sex, y = UniqueFeederCount)) +
  geom_boxplot() +
  theme_bw()

ggplot(dates, aes(x = UniqueFeederCount)) +
  geom_histogram() +
  facet_wrap(~Sex) #I used this type of graph only for this data because Sex only has two categories so it is better to visualize it like this
```
